---
title: "Soil Data Visualization Report"
author: "ebisushingo"
date: "2025-05-26"
output: html_document
---

# Introduction
This report visualizes and explores the soil dataset (`gakujitsu2025data - シート1.csv`) using R and ggplot2. The analysis focuses on the relationships between soil order, SOM color, SOM alkali, pH, and various soil chemical properties.

# Data Import
```{r setup, message=FALSE, warning=FALSE}
library(readr)
library(ggplot2)
library(tidyr)
file_path <- "gakujitsu2025data - シート1.csv"
data <- read_csv(file_path, locale = locale(encoding = "UTF-8"), show_col_types = FALSE)
if ("...1" %in% names(data)) {
  data <- data[ , !(names(data) %in% "...1")]
}
head(data)
```

# Visualization 1: SOM color and SOM alkali by USDA Soil Order
```{r plot1}
p1 <- ggplot(data, aes(x = factor(USDA_SoilOrder), y = SOM_color, color = SOM_alkali)) +
  geom_jitter(width = 0.2, size = 3) +
  labs(x = "USDA Soil Order", y = "SOM color", color = "SOM alkali", title = "SOM color and SOM alkali by USDA Soil Order") +
  theme_minimal()
print(p1)
```

# Visualization 2: pH_H2O and pH_KCl by SOM alkali
```{r plot2}
data_long <- pivot_longer(data, cols = c(pH_H2O, pH_KCl), names_to = "pH_type", values_to = "pH_value")
p2 <- ggplot(data_long, aes(x = SOM_alkali, y = pH_value, color = pH_type)) +
  geom_point(size = 3) +
  geom_line(aes(group = Name), alpha = 0.5) +
  labs(x = "SOM alkali", y = "pH value", color = "pH type", title = "Relationship between SOM alkali and pH values") +
  theme_minimal()
print(p2)
```

# Visualization 3-1: deltapH, exH, CEC_eff, CEC_pot by USDA Soil Order
```{r plot3-1}
vars1 <- c("deltapH", "exH", "CEC_eff", "CEC_pot")
data_box1 <- pivot_longer(data, cols = all_of(vars1), names_to = "variable", values_to = "value")
p3_1 <- ggplot(data_box1, aes(x = factor(USDA_SoilOrder), y = value, fill = factor(USDA_SoilOrder))) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, size = 1, alpha = 0.5) +
  facet_wrap(~variable, scales = "free_y") +
  labs(x = "USDA Soil Order", y = "Value", fill = "USDA Soil Order", title = "Comparison of deltapH, exH, CEC_eff, CEC_pot by USDA Soil Order") +
  theme_minimal()
print(p3_1)
```

# Visualization 3-2: CationSaturation, ExCa, ExMg, ExK by USDA Soil Order
```{r plot3-2}
vars2 <- c("CationSaturation", "ExCa", "ExMg", "ExK")
data_box2 <- pivot_longer(data, cols = all_of(vars2), names_to = "variable", values_to = "value")
p3_2 <- ggplot(data_box2, aes(x = factor(USDA_SoilOrder), y = value, fill = factor(USDA_SoilOrder))) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, size = 1, alpha = 0.5) +
  facet_wrap(~variable, scales = "free_y") +
  labs(x = "USDA Soil Order", y = "Value", fill = "USDA Soil Order", title = "Comparison of CationSaturation, ExCa, ExMg, ExK by USDA Soil Order") +
  theme_minimal()
print(p3_2)
```

# Visualization 3-3: ExNa, Pabsorption by USDA Soil Order
```{r plot3-3}
vars3 <- c("ExNa", "Pabsorption")
data_box3 <- pivot_longer(data, cols = all_of(vars3), names_to = "variable", values_to = "value")
p3_3 <- ggplot(data_box3, aes(x = factor(USDA_SoilOrder), y = value, fill = factor(USDA_SoilOrder))) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, size = 1, alpha = 0.5) +
  facet_wrap(~variable, scales = "free_y") +
  labs(x = "USDA Soil Order", y = "Value", fill = "USDA Soil Order", title = "Comparison of ExNa, Pabsorption by USDA Soil Order") +
  theme_minimal()
print(p3_3)
```

# Discussion
The visualizations reveal several trends in the soil data:

- **SOM color and SOM alkali**: There is a clear difference in SOM color and SOM alkali among different USDA Soil Orders. Andisol soils tend to have higher SOM alkali and darker SOM color, while Ultisol and Inceptisol show lower values.
- **pH values**: The relationship between SOM alkali and pH (both H2O and KCl) shows that higher SOM alkali is generally associated with higher pH values, but the trend varies by soil type.
- **Soil chemical properties**: Boxplots for deltapH, exH, CEC, cation saturation, and exchangeable cations show that Andisol soils generally have higher CEC and cation saturation, while other soil orders have lower values. ExNa and Pabsorption also vary by soil order, with some outliers present.

Overall, the data suggest that soil order is a strong determinant of chemical properties, and that Andisol soils in particular have distinct characteristics compared to other orders in this dataset.
