---
title: "カイコ鼓動データ解析レポート"
author: "ebisushingo"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: united
    highlight: tango
  pdf_document:
    toc: true
    latex_engine: xelatex
---

# 概要

本レポートでは、`data/sakamoto実験 - シート1.csv` のカイコ鼓動データを解析し、温度ごとの平均・標準偏差・グラフを作成します。

df <- read.csv("../data/sakamoto実験 - シート1.csv", check.names = FALSE, skip = 4, header = TRUE)
temp_cols <- setdiff(colnames(df), "個体番号")
temp_cols <- temp_cols[temp_cols != "" & !is.na(temp_cols)]
df_long <- df %>%
df_long <- df_long %>% arrange(温度)
# Data import and wrangling (Sheet 1 & 2)

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
# Sheet 1: skip 4 lines, header on 5th
df <- read.csv("../data/sakamoto実験 - シート1.csv", check.names = FALSE, skip = 4, header = TRUE)
colnames(df)[1] <- "個体番号"
temp_cols <- setdiff(colnames(df), "個体番号")
temp_cols <- temp_cols[temp_cols != "" & !is.na(temp_cols)]
df_long <- df %>%
  pivot_longer(
    all_of(temp_cols),
    names_to = "温度",
    values_to = "鼓動回数"
  ) %>%
  mutate(
    温度 = as.numeric(gsub("[^0-9.]", "", 温度)),
    鼓動回数 = as.numeric(鼓動回数)
  ) %>%
  drop_na()
df_long <- df_long %>% arrange(温度)

# Sheet 2: skip 4 lines, header on 5th
df2 <- read.csv("../data/sakamoto実験 - シート2.csv", check.names = FALSE, skip = 0, header = TRUE)
colnames(df2)[1] <- "Individual"
df2_long <- df2 %>%
  pivot_longer(
    -Individual,
    names_to = "Closed_Spiracle",
    values_to = "Beats_30sec"
  ) %>%
  mutate(
    Beats_30sec = as.numeric(Beats_30sec)
  ) %>%
  drop_na()
df2_long <- df2_long %>% arrange(Closed_Spiracle)
```


# Summary statistics by temperature (Sheet 1)

```{r summary-stats}
summary_stats <- df_long %>%
  group_by(温度) %>%
  summarise(
    n = n(),
    mean = mean(鼓動回数, na.rm = TRUE),
    sd = sd(鼓動回数, na.rm = TRUE)
  )
colnames(summary_stats) <- c("Temperature", "N", "Mean", "SD")
knitr::kable(summary_stats, caption = "Summary statistics by temperature (Sheet 1)")
```

# Summary statistics by closed spiracle (Sheet 2)

```{r summary-stats2}
summary_stats2 <- df2_long %>%
  group_by(Closed_Spiracle) %>%
  summarise(
    n = n(),
    mean = mean(Beats_30sec, na.rm = TRUE),
    sd = sd(Beats_30sec, na.rm = TRUE)
  )
knitr::kable(summary_stats2, caption = "Summary statistics by closed spiracle (Sheet 2)")
```


# Mean and SD by temperature (Sheet 1)

```{r plot-mean-sd, fig.width=7, fig.height=5}
p <- ggplot(summary_stats, aes(x = Temperature, y = Mean)) +
  geom_line(group = 1, color = "steelblue", size = 1.2) +
  geom_point(size = 3, color = "steelblue") +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width = 0.5, color = "orange") +
  labs(
    title = "Mean heartbeats per minute by temperature (Sheet 1)",
    x = "Temperature (C)",
    y = "Heartbeats per minute"
  ) +
  theme_minimal(base_size = 14)
print(p)
```

# Mean and SD by closed spiracle (Sheet 2)

```{r plot-mean-sd2, fig.width=7, fig.height=5}
p3 <- ggplot(summary_stats2, aes(x = Closed_Spiracle, y = mean, group = 1)) +
  geom_line(color = "darkred", size = 1.2) +
  geom_point(size = 3, color = "darkred") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.5, color = "orange") +
  labs(
    title = "Mean heartbeats per 30 sec by closed spiracle (Sheet 2)",
    x = "Closed spiracle",
    y = "Heartbeats per 30 sec"
  ) +
  theme_minimal(base_size = 14)
print(p3)
```


# Individual trajectories (Sheet 1)

```{r plot-individual, fig.width=7, fig.height=5}
df_long <- df_long %>%
  rename(Individual = 個体番号, Temperature = 温度, Heartbeats = 鼓動回数)
p2 <- ggplot(df_long, aes(x = Temperature, y = Heartbeats, group = Individual, color = Individual)) +
  geom_line(alpha = 0.5) +
  geom_point() +
  labs(
    title = "Individual heartbeats per minute by temperature (Sheet 1)",
    x = "Temperature (C)",
    y = "Heartbeats per minute"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
print(p2)
```

# Individual trajectories by closed spiracle (Sheet 2)

```{r plot-individual2, fig.width=7, fig.height=5}
p4 <- ggplot(df2_long, aes(x = Closed_Spiracle, y = Beats_30sec, group = Individual, color = Individual)) +
  geom_line(alpha = 0.5) +
  geom_point() +
  labs(
    title = "Individual heartbeats per 30 sec by closed spiracle (Sheet 2)",
    x = "Closed spiracle",
    y = "Heartbeats per 30 sec"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
print(p4)
```

# Discussion

- How did heartbeats change as temperature increased?
- How large was the individual variation?
- Future directions for experiments and analysis
